---
title: "p8105_hw2_yl4924"
author: "Yongzheng Li"
date: "2022-10-02"
output: github_document
---

# load the package
```{r}
library(tidyverse)
library(readxl)
```




# Problem 1

### import and clean the data

```{r}
nyc_transit_data <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",  
                    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
                    janitor::clean_names() %>%
                    select(line, station_name, station_latitude, station_longitude, 
                           starts_with("routes"), entry, vending, entrance_type,ada) %>% 
                    mutate(ifelse(entry == "Yes", TRUE, FALSE))
                    
```



# Problem 2

### Read and clean the Mr. Trash Wheel sheet
```{r}
mr_trash <- read_excel("./data/Trash Wheel Collection Data.xlsx", range = "A2:N549", sheet = "Mr. Trash Wheel" ) %>% 
            janitor::clean_names() %>% 
            mutate(
              sports_balls = round(sports_balls),
              sports_balls = as.integer(sports_balls))
```


### Read and clean the Professor Trash Wheel sheet
```{r}
professor_trash <- read_excel("./data/Trash Wheel Collection Data.xlsx", range = "A2:M96", sheet = "Professor Trash Wheel") %>% 
                   janitor::clean_names() %>% 
                   mutate(
                     sports_balls = 0,
                     sports_balls = as.integer(sports_balls)) %>% 
                   relocate(sports_balls, .after = chip_bags)
                   
```


### add additional variable 'trash_wheel_type'
after view the two datasets, I decide to use binding to combine mr. trash wheel and professor_trash wheel. Since both datasets have variable 'dumpster', I decide to create variable 'trash_wheel_type' to seperate them.
```{r}
mr_trash <- mutate(mr_trash, trash_wheel_type = "Mr")
professor_trash <- mutate(professor_trash, trash_wheel_type = "Professor")
```

### combine two dataset





