---
title: "p8105_hw2_yl4924"
author: "Yongzheng Li"
date: "2022-10-02"
output: github_document
---

# load the package
```{r}
library(tidyverse)
library(readxl)
```




# Problem 1

### import and clean the data

```{r}
nyc_transit_data <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",  
                    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
                    janitor::clean_names() %>%
                    select(line, station_name, station_latitude, station_longitude, 
                           starts_with("routes"), entry, vending, entrance_type,ada) %>% 
                    mutate(ifelse(entry == "Yes", TRUE, FALSE))
                    
```



# Problem 2

### Read and clean the Mr. Trash Wheel sheet
Variable 'year' in Mr. Trash Wheel is character variable. I decide to convert year to numerical variable.
```{r}
mr_trash <- read_excel("./data/Trash Wheel Collection Data.xlsx", range = "A2:N549", sheet = "Mr. Trash Wheel" ) %>% 
            janitor::clean_names() %>% 
            mutate(
              sports_balls = round(sports_balls),
              sports_balls = as.integer(sports_balls),
              year = as.numeric(year))
```


### Read and clean the Professor Trash Wheel sheet
```{r}
professor_trash <- read_excel("./data/Trash Wheel Collection Data.xlsx", range = "A2:M96", sheet = "Professor Trash Wheel") %>% 
                   janitor::clean_names() %>% 
                   mutate(
                     sports_balls = 0,
                     sports_balls = as.integer(sports_balls)) %>% 
                   relocate(sports_balls, .after = chip_bags)
                   
```


### add additional variable 'trash_wheel_type'
After view the two datasets, I decide to use binding to combine mr. trash wheel and professor_trash wheel. Since both datasets have variable 'dumpster', I decide to create variable 'trash_wheel_type' as the identity variable.
```{r}
mr_trash <- mutate(mr_trash, trash_wheel_type = "Mr")
professor_trash <- mutate(professor_trash, trash_wheel_type = "Professor")
```

### combine two dataset
```{r}
trash_tidy <- bind_rows(mr_trash, professor_trash)
```

The combined dataset contain `r nrow(trash_tidy)` rows and `r ncol(trash_tidy)` columns. Among 15 columns, the key variables are weight in tons and volume in cubic yards. The mean weight of trash collect of the combined data is `r mean(trash_tidy$weight_tons)` tons. The mean volume of trash collect of the combined data is `r mean(trash_tidy$volume_cubic_yards)` cubic yards. The total weight of trash collected by Professor Trash Wheel = `r sum(professor_trash$weight_tons)` tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r filter(mr_trash, year == 2020) %>% {sum(.$sports_balls)}`.


# Problem 3

### read and clean the data of pols-month.csv
```{r}
pols_month <- read_csv("./data/pols-month.csv") %>% 
              janitor::clean_names() %>% 
              separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
              mutate(
                year = as.integer(year),
                month = month.abb[as.integer(month)],
                day = as.integer(day),
                prez_gop = recode(prez_gop, `1` = "gop", `0` = "dem", `2` = "otherwise" ),
                prez_dem = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% 
              pivot_longer(
                cols = c(prez_gop, prez_dem),
                names_to = NULL,
                values_to = "president") %>% 
              select(-contains("prez")) %>% 
              select(-day) %>% 
              distinct()
              
```

### read and clean the data of snp.csv
```{r}
snp <- read_csv("./data/snp.csv") %>%
       janitor::clean_names() %>% 
       separate(date, into = c("year", "month", "day"), sep = "/") %>% 
       mutate(
         year = as.integer(year),
         day = as.integer(day),
         month = as.integer(month),
         year_before_2001 = case_when(50 <= day & day <= 99 ~ 1900+day,
                                      day == 0 ~2000+day,
                                      1 <= day & day <= 49 ~ 0),
         
         month = case_when(year_before_2001 > 0 ~ year+0,
                           year_before_2001 == 0 ~ month+0),
         month = month.abb[month],
         
         year = case_when(year_before_2001 == 0 ~ 2000+year,
                          year_before_2001 > 0 ~ 0+year_before_2001)) %>% 
      select(-c(day,year_before_2001))
head(snp)
```
I noticed the date in years after 2001(yyyy/mm/dd) is different from date in years before 2001(mm/dd/yy). 
I decide to create a new variable 'year_before_2001' used to seperate years before 2001 and after 2001. 
Then I mutate month for years before 2001(from mm/dd/yy to mm/mm/yy since we do not need day so I just override it) 
Next, I mutate years for years before 2001 based on the variable 'year_before_2001' (from mm/mm/yy to yyyy/mm/yy) and change years to 4 digit for years after 2001.
Finally, I drop the variable 'day' and variable 'year_before_2001' to keep the columns I need.

###  tidy the unemployment data
```{r}
unemployment <- read_csv("./data/unemployment.csv") %>% 
                pivot_longer(
                  cols = Jan:Dec,
                  names_to = "month",
                  values_to = "percentage") %>% 
                janitor::clean_names()
```

### join the dataset
```{r}
fte_data <- left_join(pols_month, snp, by = c("year", "month") ) %>% 
left_join(unemployment, by = c("year", "month"))
head(fte_data)
```


the data pols_month contain the number of national politicians who are democratic or republican from 1947 to 2015. Data snp contain information of closing value of stock market index from 1950 to `r max(snp$year)`. Data unemployment contain the information of percent of unemployment from `r min(unemployment$year)` to `r max(unemployment$year)`. The combined data fte_data contain `r nrow(fte_data)` rows and `r ncol(fte_data)` columns. the range of years is from 1947 to 2015. The key variables is president, close, and percentage.










